### 2021-04-20 FastLinAlg/M2/ChangesMade_ResponseToReviewers.txt
Various changes on the paper, and on the main code.

We thank the reviewers for their helpful suggestions, comments and preliminary codes and examples. We would be very happy to invite the reviewer to be a coauthor to the source code (and also for the paper, if desired).

Changes Made to Paper:

a) Section 2.1 (Page 2),

b) 2.2 Page 3;

c) Sentence before Example 2.5 added on Page 4

d) Page 5 Remark 2.2.1 added following reviewer's suggestion to discuss a bit about the probability of succesful output.

e) Page 5 Section 5 dimViaBezout added with example 3.1 and paragraph following it.

MAIN CHANGES:

-- We have implemented the dimViaBezout to compute the dimension probabilistically. This intersects V(I) with successively higher dimensional random linear spaces until there is an intersection.  For example, if $V(I)$ intersect a random line has a point, then we expect that $V(I)$ contains a hypersurface.  If there was no intersection, this function tries a 2-dimensional linear space, and so on.  This speeds up many computations.

    Some comments on implementation.  The referee suggested using a binary search pattern, however we found that in practice, it is much faster to start with a 1-dimensional linear space, then a 2-dimensional linear space, and so on, until we find an intersection.  This is because in our examples, checking whether an intersection is empty is much faster when the intersection is indeed empty.

    Furthermore, our implementation is in the affine case and not the projective setting (it did not seem to impact performance, and now it works on a wider range of examples).

Optional Inputs:

DimensionIntersectionAttempts => ZZ : the number of linear spaces to try before moving to the next dimension
MinimumFieldSize => ZZ: if the ambient field is smaller than this value it will automatically be replaced with an extension. (The user may set the  MinimumFieldSize to ensure that the field being worked over is big enough.  For instance, there are relatively few linear spaces over a field of characteristic 2, and this can cause incorrect results to be provided.)

-- The MultiplicationTable strategy has also beeen added. (This MultiplicationTable strategy essentially is still a LinearIntersection but it handles the point issue differently. The previous strategy was using decompose to break up the ideal.)

-- Following the reviewer's suggestion, we have added Remark 2.2.1 with references that all point to a good probability of finding a rational point by intersecting with a line in the case of hypersurface.

Changes Made to the Source Code:

We have added the MultiplicationStrategy verbatim as provided by the reviewer but have changed the computation of the dimension to be done via dimViaBezout instead of dimDegViaBezout. (This MultiplicationTable strategy essentially is still a LinearIntersection but it handles the point issue differently. The previous strategy was using decompose to break up the ideal.)

There are some aspects of this code we do not totally understand.  In particular, we do not understand what is occurring in the lines:

m0:=contract(transpose matrix entries b2,matrix entries((xx_0*b1)%Js));
and what follows.

Additionally, we have had quite a few examples, especially when running something iteratively, where the saturation:
Js:=saturateInGenericCoordinates J;
(which only saturates in the last coordinate)
will saturate away the points, and thus causes this function to crash.

However, doing a saturation with respect to every variable (even separately) seems to make this function quite a bit slower.  We are tempted to saturate with respect to a (random) variable, then use a try command to see if that yields a good result, and if it crashed try saturating with respect to a different (random) variable.  However, we

New documentation and examples have been provided.

All the strategies now make use of the dimViaBezout to compute the dimension of the given ideal, if the codimension is not already known. In case, that it is known, it should be specified along with the call of randomPoints.

ProjectionAttempts has been set to 0 in the Default Strategy now for reasons discussed.

-------------------------------

We also provide some examples demonstrating performance in some somewhat larger examples.

-------------------------
----Large Example #1-----
-------------------------

-------------------------
----Large Example #3-----
-------------------------

Here is an example (coming out of a jacobian computation) where the rational points tend to be saturated away, which causes the MultiplicationTable strategy to fail (at least based on our understanding).

In this example, the primary speed limitation still appears to be the dimension computation.

T2 = ZZ/101[x1,x2,x3,x4,x5,x6,x7, Z];
J2 = ideal(x5*x6-x4*x7,x1*x6-x2*x7,x5^2-x1*x7,x4*x5-x2*x7,x4^2-x2*x6,x1*x4-x2*x5,x2*x3^3*x5+3*x2*x3^2*x7*Z+8*x2^2*x5*Z^2+3*x3*
      x4*x7*Z^2-8*x4*x7*Z^3+x6*x7*Z^3,x1*x3^3*x5+3*x1*x3^2*x7*Z+8*x1*x2*x5*Z^2+3*x3*x5*x7*Z^2-8*x5*x7*Z^3+x7^2*Z^3,x2*x3^3*x4+3*
      x2*x3^2*x6*Z+8*x2^2*x4*Z^2+3*x3*x4*x6*Z^2-8*x4*x6*Z^3+x6^2*Z^3,x2^2*x3^3+3*x2*x3^2*x4*Z+8*x2^3*Z^2+3*x2*x3*x6*Z^2-8*x2*x6*
      Z^3+x4*x6*Z^3,x1*x2*x3^3+3*x2*x3^2*x5*Z+8*x1*x2^2*Z^2+3*x2*x3*x7*Z^2-8*x2*x7*Z^3+x4*x7*Z^3,x1^2*x3^3+3*x1*x3^2*x5*Z+8*x1^2
      *x2*Z^2+3*x1*x3*x7*Z^2-8*x1*x7*Z^3+x5*x7*Z^3,x1*x2^2*x4-x2^3*x5,-x1^2*x3^3*x5^2-6*x1*x3^2*x5^3*Z-6*x3*x5^4*Z^2-3*x1*x3*x5^
      2*x7*Z^2+16*x5^4*Z^3+8*x1*x5^2*x7*Z^3-4*x5^3*x7*Z^3,-x2^2*x3^3*x4^2-6*x2*x3^2*x4^3*Z-8*x2^3*x4^2*Z^2-6*x3*x4^4*Z^2-3*x2*x3
      *x4^2*x6*Z^2+16*x4^4*Z^3+8*x2*x4^2*x6*Z^3-4*x4^3*x6*Z^3,-2*x3^3*x5^2*x7^2-x1*x3^3*x7^3-6*x3^2*x5*x7^3*Z-8*x1*x5^2*x6*x7*Z^
      2-8*x1*x4*x5*x7^2*Z^2-16*x2*x5^2*x7^2*Z^2-8*x1*x2*x7^3*Z^2-3*x3*x7^4*Z^2+8*x7^4*Z^3,-8*x1^5,-x1^2*x3^3*x5^2+x1^3*x3^3*x7-6
      *x1*x3^2*x5^3*Z+6*x1^2*x3^2*x5*x7*Z+8*x1^3*x4*x5*Z^2-16*x1^2*x2*x5^2*Z^2-6*x3*x5^4*Z^2+8*x1^3*x2*x7*Z^2+3*x1*x3*x5^2*x7*Z^
      2+3*x1^2*x3*x7^2*Z^2+16*x5^4*Z^3-8*x1*x5^2*x7*Z^3-4*x5^3*x7*Z^3-8*x1^2*x7^2*Z^3+4*x1*x5*x7^2*Z^3,-x1^2*x5^2+x1^3*x7,-6*x1^
      2*x2*x3^5*x5*x7-30*x1*x2*x3^4*x5^2*x7*Z-48*x1^2*x2^2*x3^2*x5*x7*Z^2-36*x2*x3^3*x5^3*x7*Z^2-24*x1*x2*x3^3*x5*x7^2*Z^2+5*x1*
      x2^2*x3*x5^2*x7*Z^3+48*x1*x2*x3^2*x5*x7^2*Z^3+47*x2*x3^2*x5^2*x7^2*Z^3-3*x1*x2*x3^2*x7^3*Z^3-48*x1*x2^2*x5*x7^2*Z^4-5*x2*
      x3*x5^2*x7^2*Z^4-24*x2*x3*x5*x7^3*Z^4+48*x2*x5*x7^3*Z^5-3*x2*x7^4*Z^5,18*x1^2*x2^3*x3^6+24*x1^2*x2^3*x3^5*Z+24*x1^2*x2^2*
      x3^5*x4*Z+9*x1*x2^3*x3^5*x5*Z+29*x1^2*x2^4*x3^3*Z^2+29*x1^2*x2^2*x3^4*x4*Z^2+19*x1*x2^3*x3^4*x5*Z^2-6*x1*x2^2*x3^4*x4*x5*Z
      ^2-9*x2^3*x3^4*x5^2*Z^2+18*x1*x2^3*x3^4*x7*Z^2-10*x1^2*x2^4*x3^2*Z^3-24*x1^2*x2^3*x3^2*x4*Z^3-43*x1*x2^4*x3^2*x5*Z^3-45*x1
      *x2*x3^3*x4^2*x5*Z^3-18*x2^2*x3^3*x4*x5^2*Z^3+43*x1^2*x2^2*x3^3*x6*Z^3+9*x1*x2^2*x3^3*x5*x6*Z^3+24*x1*x2^3*x3^3*x7*Z^3+24*
      x1*x2^2*x3^3*x4*x7*Z^3-20*x1*x2^4*x3*x5*Z^4-48*x1*x2^3*x3*x4*x5*Z^4-39*x1*x2*x3^2*x4^2*x5*Z^4-15*x2*x3^2*x4^2*x5^2*Z^4+10*
      x1^2*x2^2*x3^2*x6*Z^4+19*x1*x2^2*x3^2*x5*x6*Z^4-18*x1*x2*x3^2*x4*x5*x6*Z^4+3*x2^2*x3^2*x5^2*x6*Z^4+6*x1^2*x2*x3^2*x6^2*Z^4
      +29*x1*x2^4*x3*x7*Z^4+29*x1*x2^2*x3^2*x4*x7*Z^4+20*x1*x2*x3*x4^2*x5*Z^5+48*x2*x3*x4^2*x5^2*Z^5+48*x1*x2*x3*x4*x5*x6*Z^5+9*
      x1*x3*x4^2*x5*x6*Z^5-48*x2^2*x3*x5^2*x6*Z^5-6*x2*x3*x4*x5^2*x6*Z^5-6*x1*x2*x3*x5*x6^2*Z^5-10*x1*x2^4*x7*Z^5-24*x1*x2^3*x4*
      x7*Z^5+43*x1*x2^2*x3*x6*x7*Z^5-24*x1*x4^2*x5*x6*Z^6+3*x4^2*x5^2*x6*Z^6+48*x1*x2*x5*x6^2*Z^6-6*x2*x5^2*x6^2*Z^6+10*x1*x2^2*
      x6*x7*Z^6+6*x1*x2*x6^2*x7*Z^6,-x1^3*x4+x1^2*x2*x5,2*x1*x2*x4^2-2*x2^2*x4*x5-x1*x2^2*x6+x2^3*x7,-2*x3^3*x5^2*x7^2-x1*x3^3*
      x7^3-6*x3^2*x5*x7^3*Z-16*x1*x5^2*x6*x7*Z^2-16*x2*x5^2*x7^2*Z^2-8*x1*x2*x7^3*Z^2-3*x3*x7^4*Z^2+8*x7^4*Z^3,-2*x4^4+2*x2*x4^2
      *x6,-27*x1*x2^3*x3^5*x5-29*x1*x2^3*x3^4*x5*Z-36*x1*x2^2*x3^4*x4*x5*Z-27*x2^3*x3^4*x5^2*Z+43*x1*x2^2*x3^3*x4*x5*Z^2-9*x1*x2
      *x3^3*x4^2*x5*Z^2+47*x2^2*x3^3*x4*x5^2*Z^2-9*x1*x2^2*x3^3*x5*x6*Z^2+18*x2^3*x3^3*x5*x7*Z^2+10*x1*x2^2*x3^2*x4*x5*Z^3+24*x1
      *x2*x3^2*x4^2*x5*Z^3-10*x2^3*x3^2*x5^2*Z^3+38*x2^2*x3^2*x4*x5^2*Z^3-15*x2*x3^2*x4^2*x5^2*Z^3+24*x1*x2^2*x3^2*x5*x6*Z^3-3*
      x1*x2*x3^2*x4*x5*x6*Z^3-18*x2^2*x3^2*x5^2*x6*Z^3+10*x2^3*x3^2*x5*x7*Z^3+6*x2^2*x3^2*x4*x5*x7*Z^3+20*x2^2*x3*x4*x5^2*Z^4+48
      *x2*x3*x4^2*x5^2*Z^4+48*x2^2*x3*x5^2*x6*Z^4-6*x2*x3*x4*x5^2*x6*Z^4-20*x2^3*x3*x5*x7*Z^4-48*x2^2*x3*x4*x5*x7*Z^4+9*x2*x3*x4
      ^2*x5*x7*Z^4-18*x2^2*x3*x5*x6*x7*Z^4-24*x2*x4^2*x5*x7*Z^5+3*x4^3*x5*x7*Z^5+48*x2^2*x5*x6*x7*Z^5-6*x2*x4*x5*x6*x7*Z^5,-3*x1
      ^2*x2*x3^5*x4*x5-6*x1*x2^2*x3^5*x5^2-12*x1*x2*x3^4*x4*x5^2*Z-9*x2^2*x3^4*x5^3*Z-12*x1^2*x2*x3^4*x4*x7*Z-12*x1*x2^2*x3^4*x5
      *x7*Z-24*x1^2*x2^2*x3^2*x4*x5*Z^2-48*x1*x2^3*x3^2*x5^2*Z^2-15*x1^2*x3^3*x4^2*x7*Z^2-48*x1*x2*x3^3*x4*x5*x7*Z^2-27*x2^2*x3^
      3*x5^2*x7*Z^2+48*x1*x2^2*x3*x4*x5^2*Z^3+5*x1^2*x2^2*x3*x4*x7*Z^3+24*x1^2*x3^2*x4^2*x7*Z^3+5*x1*x2^3*x3*x5*x7*Z^3+24*x1*x2*
      x3^2*x4*x5*x7*Z^3-36*x1*x3^2*x4^2*x5*x7*Z^3+24*x2^2*x3^2*x5^2*x7*Z^3-12*x2*x3^2*x4*x5^2*x7*Z^3-18*x1*x2*x3^2*x4*x7^2*Z^3-
      18*x2^2*x3^2*x5*x7^2*Z^3-48*x1^2*x2*x4^2*x7*Z^4-24*x1*x2^2*x4*x5*x7*Z^4+48*x1*x3*x4^2*x5*x7*Z^4+48*x1*x2*x3*x4*x7^2*Z^4-18
      *x1*x3*x4^2*x7^2*Z^4+48*x2^2*x3*x5*x7^2*Z^4-15*x2*x3*x4*x5*x7^2*Z^4+48*x1*x4^2*x7^2*Z^5+24*x2*x4*x5*x7^2*Z^5-3*x4^2*x5*x7^
      2*Z^5,0,9*x1^3*x2^2*x3^7+9*x1^3*x2*x3^6*x4*Z+36*x1^2*x2^2*x3^6*x5*Z-29*x1^3*x2^3*x3^4*Z^2-24*x1^3*x2*x3^5*x4*Z^2+24*x1^2*
      x2^2*x3^5*x5*Z^2+39*x1^2*x2*x3^5*x4*x5*Z^2+30*x1*x2^2*x3^5*x5^2*Z^2+3*x1^3*x2*x3^5*x6*Z^2+18*x1^2*x2^2*x3^5*x7*Z^2-29*x1^3
      *x2^2*x3^3*x4*Z^3-29*x1^2*x2^3*x3^3*x5*Z^3-19*x1^2*x2*x3^4*x4*x5*Z^3-29*x1*x2^2*x3^4*x5^2*Z^3+45*x1*x2*x3^4*x4*x5^2*Z^3+15
      *x1^2*x2*x3^4*x5*x6*Z^3-24*x1^2*x2^2*x3^4*x7*Z^3+24*x1^2*x2*x3^4*x4*x7*Z^3+6*x1*x2^2*x3^4*x5*x7*Z^3+10*x1^3*x2^2*x3^2*x4*Z
      ^4-10*x1^2*x2^3*x3^2*x5*Z^4-5*x1^2*x2^2*x3^2*x4*x5*Z^4+24*x1*x2^3*x3^2*x5^2*Z^4-43*x1*x2*x3^3*x4*x5^2*Z^4+24*x1^3*x2^2*x3^
      2*x6*Z^4+18*x1*x2*x3^3*x5^2*x6*Z^4+29*x1^2*x2^3*x3^2*x7*Z^4-43*x1^2*x2*x3^3*x4*x7*Z^4-6*x1^2*x3^3*x4^2*x7*Z^4+43*x1*x2^2*
      x3^3*x5*x7*Z^4+48*x1*x2*x3^3*x4*x5*x7*Z^4+12*x1^2*x2*x3^3*x6*x7*Z^4-27*x1*x2^2*x3^3*x7^2*Z^4+20*x1^2*x2^2*x3*x4*x5*Z^5+48*
      x1^2*x2^2*x3*x5*x6*Z^5-20*x1^2*x2^3*x3*x7*Z^5+5*x1^2*x2^2*x3*x4*x7*Z^5-10*x1^2*x2*x3^2*x4*x7*Z^5+48*x1*x2^3*x3*x5*x7*Z^5+
      10*x1*x2^2*x3^2*x5*x7*Z^5+39*x1*x2*x3^2*x4*x5*x7*Z^5-9*x1*x3^2*x4^2*x5*x7*Z^5+15*x2*x3^2*x4*x5^2*x7*Z^5-24*x1^2*x2*x3^2*x6
      *x7*Z^5+3*x1^2*x3^2*x4*x6*x7*Z^5+27*x1*x2*x3^2*x5*x6*x7*Z^5+14*x1*x2^2*x3^2*x7^2*Z^5-27*x1*x2*x3^2*x4*x7^2*Z^5-48*x1^2*x2*
      x4^2*x7*Z^6+24*x1*x2^2*x4*x5*x7*Z^6-20*x1*x2*x3*x4*x5*x7*Z^6-48*x2*x3*x4*x5^2*x7*Z^6+24*x1^2*x2^2*x6*x7*Z^6-48*x1*x2*x3*x5
      *x6*x7*Z^6+6*x1*x3*x4*x5*x6*x7*Z^6+20*x1*x2^2*x3*x7^2*Z^6-5*x1*x2*x3*x4*x7^2*Z^6-18*x1*x3*x4^2*x7^2*Z^6+3*x2*x3*x4*x5*x7^2
      *Z^6+9*x1*x2*x3*x6*x7^2*Z^6+48*x1*x4^2*x7^2*Z^7-24*x2*x4*x5*x7^2*Z^7-3*x4^2*x5*x7^2*Z^7-24*x1*x2*x6*x7^2*Z^7+3*x1*x4*x6*x7
      ^2*Z^7,3*x1*x3^5*x4*x5^2+3*x1*x2*x3^5*x5*x7+8*x1^2*x2*x3^3*x4*x5*Z+9*x1*x3^4*x5^2*x6*Z+9*x1*x2*x3^4*x7^2*Z+48*x1*x2*x3^2*
      x4*x5^2*Z^2+24*x1^2*x2*x3^2*x5*x6*Z^2+24*x1^2*x2*x3^2*x4*x7*Z^2+24*x1*x2^2*x3^2*x5*x7*Z^2+x3^3*x4*x5^2*x7*Z^2+9*x1*x3^3*x5
      *x6*x7*Z^2+9*x1*x3^3*x4*x7^2*Z^2+x2*x3^3*x5*x7^2*Z^2-37*x1^2*x2^2*x4*x5*Z^3+48*x1^2*x3*x4*x5*x6*Z^3+24*x1^2*x3*x4^2*x7*Z^3
      -24*x1*x3^2*x5*x6*x7*Z^3+3*x3^2*x5^2*x6*x7*Z^3-24*x1*x3^2*x4*x7^2*Z^3+6*x1*x3^2*x6*x7^2*Z^3+3*x2*x3^2*x7^3*Z^3-27*x1^2*x4*
      x5*x6*Z^4+16*x1^2*x5*x6^2*Z^4+37*x1^2*x4^2*x7*Z^4+16*x2*x4*x5^2*x7*Z^4+16*x1^2*x4*x6*x7*Z^4+8*x2^2*x5*x7^2*Z^4+3*x3*x5*x6*
      x7^2*Z^4+3*x3*x4*x7^3*Z^4-8*x5*x6*x7^2*Z^5-8*x4*x7^3*Z^5+2*x6*x7^3*Z^5,-2*x5^4+2*x1*x5^2*x7,-2*x1*x2^3*x3^3-3*x1*x2^2*x3^2
      *x4*Z-24*x1*x2^4*Z^2+x1*x2*x4*x6*Z^3,-x2^2*x4*x5+x1*x2^2*x6,6*x1*x2^3*x3^7+8*x1*x2^3*x3^6*Z+8*x1*x2^2*x3^6*x4*Z+24*x1*x2^4
      *x3^4*Z^2-24*x1*x2^2*x3^5*x4*Z^2-18*x2^3*x3^5*x7*Z^2+27*x1*x2^4*x3^3*Z^3-45*x1*x2^3*x3^3*x4*Z^3+48*x1*x2^2*x3^4*x6*Z^3+24*
      x2^3*x3^4*x7*Z^3-30*x2^2*x3^4*x4*x7*Z^3+10*x1*x2^5*x3*Z^4+10*x1*x2^3*x3^2*x4*Z^4+37*x1*x2^2*x3^3*x6*Z^4+2*x1*x2*x3^3*x6^2*
      Z^4-29*x2^4*x3^2*x7*Z^4-37*x2^3*x3^3*x7*Z^4+27*x2^2*x3^3*x4*x7*Z^4-8*x2*x3^3*x4^2*x7*Z^4+7*x1*x2^5*Z^5+37*x1*x2^4*x4*Z^5-
      20*x1*x2^3*x3*x6*Z^5+20*x2^4*x3*x7*Z^5+48*x2^3*x3*x4*x7*Z^5+10*x2^2*x3^2*x4*x7*Z^5+24*x2*x3^2*x4^2*x7*Z^5-43*x2^2*x3^2*x6*
      x7*Z^5-7*x1*x2^3*x6*Z^6+16*x1*x2^2*x6^2*Z^6+7*x2^4*x7*Z^6-27*x2^3*x4*x7*Z^6+8*x2^2*x4^2*x7*Z^6-30*x2^2*x3*x6*x7*Z^6-48*x2*
      x3*x4*x6*x7*Z^6-6*x2*x3*x6^2*x7*Z^6-7*x2^2*x6*x7*Z^7-37*x2*x4*x6*x7*Z^7+16*x2*x6^2*x7*Z^7-2*x4*x6^2*x7*Z^7,-2*x1*x2*x3^6*
      x5*x7-3*x2*x3^5*x5^2*x7*Z-48*x1*x2^2*x3^3*x5*x7*Z^2-3*x1*x3^4*x5*x6*x7*Z^2-3*x2*x3^4*x5*x7^2*Z^2-48*x2^2*x3^2*x5^2*x7*Z^3+
      24*x1^2*x2*x3^2*x6*x7*Z^3+8*x1*x3^3*x5*x6*x7*Z^3-x3^3*x5^2*x6*x7*Z^3+8*x2*x3^3*x5*x7^2*Z^3-7*x1*x3^3*x6*x7^2*Z^3+47*x1*x2^
      3*x5*x7*Z^4-48*x1*x2*x3*x5*x6*x7*Z^4-48*x2^2*x3*x5*x7^2*Z^4+24*x1*x3^2*x6*x7^2*Z^4+27*x1*x2*x5*x6*x7*Z^5-16*x2*x5^2*x6*x7*
      Z^5+8*x1^2*x6^2*x7*Z^5+27*x2^2*x5*x7^2*Z^5+16*x1*x2*x6*x7^2*Z^5+3*x3*x6*x7^3*Z^5-8*x6*x7^3*Z^6,-3*x1^2*x2^2*x3^2*x7-6*x1*
      x2^2*x3*x5*x7*Z-3*x1*x2^2*x7^2*Z^2,-6*x1*x2^2*x3^5*x5^2-6*x1*x2*x3^4*x4*x5^2*Z-6*x2^2*x3^4*x5^3*Z-18*x1*x2^2*x3^4*x5*x7*Z-
      48*x1*x2^3*x3^2*x5^2*Z^2-12*x1*x2*x3^3*x5^2*x6*Z^2-30*x2^2*x3^3*x5^2*x7*Z^2-10*x1*x2^2*x3*x4*x5^2*Z^3+15*x2^3*x3*x5^3*Z^3+
      48*x1*x2*x3^2*x5^2*x6*Z^3-18*x2*x3^2*x5^3*x6*Z^3-48*x2^2*x3^2*x5^2*x7*Z^3+18*x2*x3^2*x4*x5^2*x7*Z^3-6*x1*x2*x3^2*x5*x6*x7*
      Z^3-12*x2^2*x3^2*x5*x7^2*Z^3-5*x1*x2^2*x5^2*x6*Z^4-5*x2*x3*x5^3*x6*Z^4-43*x2^3*x5^2*x7*Z^4+5*x2*x3*x4*x5^2*x7*Z^4-12*x2*x3
      *x5^2*x6*x7*Z^4+12*x2*x3*x4*x5*x7^2*Z^4);

i55 :  time randomPoints(J2, Strategy=>MultiplicationTable)
RandomRationalPoints.m2:840:17:(3):[12]: error: expected a polynomial ring
RandomRationalPoints.m2:840:17:(3):[12]: --entering debugger (type help to see debugger commands)                                
RandomRationalPoints.m2:840:17-840:79: --source code:                                                                            
            m0:=contract(transpose matrix entries b2,matrix entries((xx_0*b1)%Js));

i56 : time randomPoints(J2, Strategy=>Default)
     -- used 4.2498 seconds

o56 = {{0, 0, 23, 0, 0, 42, 29, 0}}

o56 : List